Identifying Heterogeneity in SAR Data with New Test Statistics

### 2024-05-02 Text/Identifying Heterogeneity in SAR Data with New Test Statistics.Rmd
Grammarly edited

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento, A.D.C.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. 
  The physical modeling of this kind of data allows to use the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell, and none dominates the return. 
  Such areas are often called "homogeneous" or  "textureless" regions. 
  The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous regions, i.e., areas in which the fully-developed speckle hypotheses do not hold. 
  We propose three test statistics to distinguish between homogeneous and inhomogeneous regions, i.e., between gamma and \(\mathcal{G}_I^0\) distributed data, both with a known number of looks. 
  The first test statistic uses a bootstrapped non-parametric estimator of Shannon entropy, which provides a robust assessment in the face of uncertain distributional assumptions.
  The second test uses the classical coefficient of variation (CV). 
  The third test uses an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. 
  We apply our test statistic to create maps of \(p\) values for the homogeneity hypothesis.
  Finally, we show that our proposal, the entropy-based test, outperforms existing methods such as the classical CV and its alternative variant, in identifying heterogeneity when applied to both simulated and actual data.
#These maps identify different types of targets.
# We show that our proposal outperforms its competitors using simulated and SAR data.
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  The author J.A. would like to thank the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and the Fundação de Amparo à Ciência e Tecnologia de Pernambuco (FACEPE) for their support. 
authorcontributions: |
  All authors (A.C.F., J.A, and A.D.C.N.) discussed the results and contributed to the research. All authors have read and agreed to the published version of the manuscript.

funding: |
  This research received no external funding.
institutionalreview: |
 Not applicable.
dataavailability: |
  The Sentinel-1  images are made publicly available by European Space Agency via Copernicus Data Space Ecosystem at \url{https://dataspace.copernicus.eu/} (accessed on 24 April 2024).
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article was written in Rmarkdown and is fully reproducible. The code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 30 April 2024).

@@

Synthetic Aperture Radar (SAR) technology has become an important tool for environmental monitoring and disaster management. It provides valuable images under various conditions, including day or night and different weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective use of SAR data depends on a thorough understanding of their statistical properties because they are corrupted by speckle.
This noise-like interference effect is inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

Speckle in intensity format is non-Gaussian, thus SAR data require reliable statistical models for accurate processing. 
The $\mathcal{G}^0$ distribution, which is suitable for SAR data, includes the
Gamma law as the limiting case for fully-developed speckle&nbsp;\cite{Ferreira2020} and provides flexibility with fewer parameters for analysis.

Our work aims to improve the identification of potential roughness features in SAR intensity data.
Physical modeling of SAR data allows the use of the Gamma distribution in the presence of fully-developed speckle, where an infinite number of independent backscatterers per resolution unit is assumed, commonly referred to as homogeneous regions

In this context, we present a set of three novel test statistics that aim to distinguish between homogeneous and non-homogeneous return, in particular between gamma and \(\mathcal{G}^0\) distributed data, assuming the number of looks is known. 
We use properties such as entropy and coefficient of variation.

Entropy is a fundamental concept in information theory with far-reaching applications in pattern recognition, statistical physics, image processing, edge detection and SAR image analysis&nbsp;\cite{Presse2013,MohammadDjafari2015,Avval2021, Nascimento2014,Nascimento2019}. 
Shannon introduced it in 1948&nbsp;\cite{Shannon1948} for a random variable as a measure of information and uncertainty.
In statistics, Shannon entropy is a crucial descriptive parameter, especially for evaluating data dispersion and performing tests for normality, exponentiality and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation is challenging, especially when the model is unknown. In these cases, non-parametric methods are used.
Spacing methods have been discussed as a non-parametric approach in Refs.&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. 
This strategy is flexible and robust because it does not enforce a model or any parametric constraints.

The coefficient of variation (CV), introduced in 1896 by Pearson&nbsp;\cite{Pearson1896}, is a relative dispersion measure widely used in various fields of applied statistics, including sampling, biostatistics, medical and biological research, climatology and other fields&nbsp;\cite{hendricks1936sampling,Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is particularly valuable for relating variables with different units. 
This is because when the main purpose is to compare the variations of several variables, the standard deviation can only serve as an adequate measure of variation if all variables are expressed in the same unit of measurement and have identical means. 
If these conditions are not met, then the CV is the relative measure that is usually used in real applications. 
The variable with the highest CV value is the one with the largest relative dispersion around the mean value&nbsp;\cite{Banik2011}.
The coefficient of variation is the basic measure of heterogeneity in SAR data&nbsp;\cite{Ulaby1986,Touzi1988}.
We study two ways of estimating the coefficient of variation.

The other parameter we study is the Shannon entropy.
Different levels of roughness, materialized as models for SAR data, have different entropy values, but this fundamental quantity can be also estimated in a model-agnostic way.
We exploit this property and design a bootstrap-improved non-parametric estimator for the Shannon entropy.

We devise test statistics based on these three estimators: 
the classical coefficient of variation, 
a robust version, 
and the Shannon entropy estimator.
We apply these test statistics to generate maps of evidence of homogeneity
that reveal different types of targets in the SAR data. 
We show that our proposed method is superior over existing approaches with simulated data and SAR images.

The article is structured as follows: Section&nbsp;\ref{sec:Background}
deals with statistical modeling and entropy estimation for intensity SAR
data. 
Section&nbsp;\ref{sec:test} outlines hypothesis tests based on non-parametric entropy and coefficients of variation estimators. 
In Section&nbsp;\ref{sec:Results} we present experimental results. 
Finally, we draw our conclusions in Section&nbsp;\ref{sec:conclusion}.

@@

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully-developed speckle and is a limiting case of the second model, which is interesting due to its versatility in accurately representing regions with different roughness properties&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density
functions (pdfs): 
\begin{align}
    f_Z(z;L, \mu\mid \Gamma_{\text{SAR}})&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z)\label{E:gamma1}\\
    \intertext{ and }
    f_Z(z; \alpha, \gamma, L \mid \mathcal{G}_I^0)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} 
where \(\mu > 0\) is the mean, 
\(\gamma > 0\) is the scale, 
\(\alpha < 0\) measures the roughness, 
\(L \geq 1\) is the number of looks, 
\(\Gamma(\cdot)\) is
the gamma function, 
and \(\mathbbm 1_{A}(z)\) is the indicator function
of the set \(A\).

@@

One of the first non-parametric estimators based on spacings was introduced by Vasicek&nbsp;\cite{vasicek1976test}. 
Under the assumption that \(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) is a random sample from the distribution \(F(z)\), the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, 
$Z_{(1)}\leq Z_{(2)}\leq\dots\leq Z_{(n)}$ are the order statistics,
and $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing, in which $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* Correa&nbsp;\cite{correa1995new}:
\begin{equation}
\widehat{H}_{\text{C}}(\bm{Z})=-\frac{1}{n} \sum_{i=1}^n \log \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\widebar{Z}_{(i)}\right)}{n\sum_{j=i-m}^{i+m}\left(Z_{(j)}-\widebar{Z}_{(i)}\right)^2},
\end{equation}
where  $\widebar{Z}_{(i)}=(2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}$, $m< \frac{n}{2}$, $Z_{(i)}=Z_{(1)}$ for $i<1$ and $Z_{(i)}=Z_{(n)}$ for $i>n$. Based on simulation, he showed that his estimator has a smaller mean square error in comparison with Vasicek's approach.

@@

These estimators are asymptotically consistent, 
but we will use improved bootstrap-improved versions because we need them to perform well with small samples.

@@

We aim at testing the following hypotheses: 
$$
\begin{cases}
  \mathcal{H}_0: \text{ The data come from the } \Gamma_{\text{SAR}}\text{ law},\\ 
  \mathcal{H}_1:\text{ The data come from the } G_I^0 \text{ distribution}.
\end{cases}
$$
We are testing the hypothesis that the data are fully-developed speckle versus the alternative of data with roughness.
As for the parametric problem, once it is not possible to define the hypothesis $\mathcal{H}_0=\alpha=-\infty$, it is impossible to solve this problem with parametric  inference alternatives (such as likelihood ratio, score, gradient and Wald hypothesis test).
The proposed tests to solve this physical problem in SAR systems are described below.

@@

This test statistic  aims to assess the behavior of the data under the null hypothesis using the empirical distribution.
If the data represent fully-developed speckle, the density should center around zero, i.e., $S(\bm{Z};L)\approx 0$. Otherwise, under the alternative hypothesis, the empirical distribution would shift from zero, suggesting significant differences and implying the presence of heterogeneous clutter.

@@

The processing time is an important feature, especially considering the application of these estimators to large datasets of SAR images, as can be seen in Section&nbsp;\ref{sec:Results}.

@@

As visible from Table&nbsp;\ref{tab:table_time}, the accuracy of the test results across the three estimators shows similarities in certain sample sizes. 
However, practical scenarios in SAR image processing often involve small sample sizes, typically obtained over windows of size $7\times7$.  
It is also noteworthy that the \(\widetilde{H}_{\text{AO}}\) estimator exhibited the shortest processing time, followed by \(\widetilde{H}_{\text{E}}\) and \(\widetilde{H}_{\text{C}}\). Considering this aspect, we select the \(\widetilde{H}_{\text{AO}}\) estimator for subsequent simulations. Henceforth,  the test statistical &nbsp;\eqref{Eq:test_e} will be denoted as: \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\).

We now verify the normality of the data generated by the \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\) test.
Figure&nbsp;\ref{fig:Plot_density} shows the empirical densities obtained by applying  the $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$ test to different sample sizes drawn from the $\Gamma_{\text{SAR}}$ distribution, where $L$ takes values $\left\{3,5, 8,11\right\}$ and $\mu=1$. Additionally, Table&nbsp;\ref{tab:table_stat_combined} summarizes the main descriptive statistics, including mean, 
standard deviation&nbsp;(SD), 
variance&nbsp;(Var), 
skewness&nbsp;(SK), 
excessive kurtosis&nbsp;(EK) and 
Anderson--Darling \(p\) values for normality. 
Results with \(p\) values greater than \(0.05\) do not indicate a violation of the normality assumption.
A low variance as well as a skewness and an excessive kurtosis of almost zero indicate limited dispersion, asymmetry and a light tail.
Normal QQ plots confirm that there is no evidence against a normal distribution, as shown in Figure&nbsp;\ref{fig:Plot_normality_qq}.

@@

After checking the normality of the data, we examined the possibilities of the proposed test in terms of size and power.
Under \(\mathcal{H}_0\), the distribution of the test statistic is asymptotically normal.
Therefore, the \(p\) values are calculated as
\(2\Phi(-|\varepsilon|)\), where \(\Phi\) is the standard
Gaussian cumulative distribution function, and \(\varepsilon\) is the standardized test statistic given by:
\[
\varepsilon=\frac{\widetilde{H}_{\text{AO}}(\bm{Z})-\left[H_{\Gamma_{\text{SAR}}}(L)+\ln \widebar{\bm{Z}}\right]}{\widehat\sigma}.
\]

We have nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and
\SI{10}{\percent}. 
In terms of size, \(1000\) simulations were used for different sample sizes from the $\Gamma_{\text{SAR}}$ distribution, with varying values of $L$, and  $\mu=1$. 
In all cases, the nominal level was achieved. 
We assessed the test power using \(1000\) simulations for different sample sizes from the $G_I^0$ distribution, with  $\mu=1$, and $\alpha=-2$.
The power generally improves with increasing sample size and number of looks.
The results are shown in Table&nbsp;\ref{tab:table_size_power}.
<!-- ACF Please clarify which alternatives you used for the power -->
<!-- JA Corrected -->

@@

In addition to the \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\) test, we also propose a test statistic based on the classical CV. This test statistic is defined as follows:
\begin{align}
	T_{\text{CV}}=\frac{S}{\widebar{Z}},
\end{align}
where $S$ and $\widebar{Z}$ are the sample standard deviation and the sample mean, respectively.

@@

We used the Akaike information criterion (AIC) and Bayesian information criterion (BIC) to select the best fitting distribution.

The AIC deals with the trade-off between the goodness-of-fit and the simplicity of the model in terms of the number of model parameters&nbsp;\cite{Burnham2004}. 
The model or distribution with the lowest value of AIC is chosen to be the best.
The BIC assesses goodness-of-fit of a distribution or model, but avoids overfitting by penalising additional degrees of freedom&nbsp;\cite{Dziak2019}. 
The model with the lowest BIC value is chosen as the best.

The AIC and BIC results in Tables&nbsp;\ref{tab:table_aic_gamma}--\ref{tab:table_aic_gio_MnADmedian} indicate that the CV and \(\text{CV}_{\text{MnAD}}\) data from different distributed $\Gamma_{\text{SAR}}$ and \(G_I^0\) synthetic sample sizes match the properties of an LN distribution.
It is important to note that this conclusion was drawn empirically based on a dictionary of distributions that are analytically tractable and well-defined under biparametric, unimodal, asymmetric, and positive distributions.

Figures&nbsp;\ref{fig:Plot_cv}--\ref{fig:Plot_madmed_gi0_MnADmedian} show empirical and fitted density plots, Q-Q plots, P-P plots as well as empirical and fitted cumulative distribution functions.
They provide qualitative sources that confirm that the LN distribution is the most appropriate distribution.

@@

This section presents the simulations we performed to evaluate the performance of the proposed test statistics, followed by applications to SAR data.

@@

The $\alpha$ parameter of the $G_I^0$ distribution is essential for interpreting texture characteristics. 
Values close to zero greater than $-3$ suggest extremely textured targets, such as urban zones&nbsp;\cite{Frery2019a}. 
As the value decreases, it indicates regions with moderate texture (in the $\left[-6,-3\right]$ region), related to forest zones, while values below $-6$ correspond to textureless regions, such as pasture, agricultural fields, and water bodies&nbsp;\cite{Neto2023}.

@@

We evaluated the proposed test statistics using three SAR images: one of the coast of Jalisco, Mexico (with a spatial resolution of 20 m both along azimuth and range directions) and two of Illinois, USA (with a spatial resolution of 10 m both along azimuth and range directions),  acquired by the Sentinel-1B satellite operating in C-band, with VV polarization and intensity format.
The first two images have a size of \(512 \times 512\) pixels, while the third has \(1024 \times 1024\) pixels  and they contain mountainous areas, agricultural regions, water bodies and  urban areas, as shown in  Figures&nbsp;\ref{fig:real_SAR_Images_coe}(a)--(c).

@@

In Figures&nbsp;\ref{fig:Mexico_crops_0.05}, \ref{fig:lake_0.05} and&nbsp;\ref{fig:Illinois_crops_0.05}, the maps of $p$-values composed of a linear gradient of black and white colors, represent the decisions at a \SI{5}{\percent} significance level. 
Values below  $0.05$  are represented by dark areas, indicating evidence to reject the null hypothesis and suggesting heterogeneity in these regions. 
In contrast, values above 0.05 are represented as white areas, indicating no evidence to reject the fully-developed speckle hypothesis.

>>>>

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento, A.D.C.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. 
  The physical modeling of this kind of data allows the use of the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell, and none dominates the return. 
  Such areas are often called "homogeneous" or  "textureless" regions. 
  The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous regions, i.e., areas where the fully-developed speckle hypotheses do not hold. 
  We propose three test statistics to distinguish between homogeneous and inhomogeneous regions, i.e., between gamma and \(\mathcal{G}_I^0\) distributed data, both with a known number of looks. 
  The first test statistic uses a bootstrapped non-parametric estimator of Shannon entropy, providing a robust assessment in uncertain distributional assumptions.
  The second test uses the classical coefficient of variation (CV). 
  The third test uses an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. 
  We apply our test statistic to create maps of \(p\) values for the homogeneity hypothesis.
  Finally, we show that our proposal, the entropy-based test, outperforms existing methods, such as the classical CV and its alternative variant, in identifying heterogeneity when applied to both simulated and actual data.
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  The author J.A. would like to thank the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and the Fundação de Amparo à Ciência e Tecnologia de Pernambuco (FACEPE) for their support. 
authorcontributions: |
  All authors (A.C.F., J.A, and A.D.C.N.) discussed the results and contributed to the research. All authors have read and agreed to the published version of the manuscript.

funding: |
  This research received no external funding.
institutionalreview: |
 Not applicable.
dataavailability: |
  The Sentinel-1  images are made publicly available by the European Space Agency via Copernicus Data Space Ecosystem at \url{https://dataspace.copernicus.eu/} (accessed on 24 April 2024).
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article was written in Rmarkdown and is fully reproducible. The code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 30 April 2024).

@@

Synthetic Aperture Radar (SAR) technology has become essential for environmental monitoring and disaster management. It provides valuable images under various conditions, including day or night and weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective use of SAR data depends on a thorough understanding of its statistical properties because it is corrupted by speckle.
This noise-like interference effect is inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

Speckle in intensity format is non-Gaussian. Thus, SAR data require reliable statistical models for accurate processing. 
The $\mathcal{G}^0$ distribution, which is suitable for SAR data, includes the
Gamma law as the limiting case for fully-developed speckle&nbsp;\cite{Ferreira2020} and provides flexibility with fewer parameters for analysis.

Our work aims to improve the identification of potential roughness features in SAR intensity data.
Physical modeling of SAR data allows the use of the Gamma distribution in the presence of fully-developed speckle, where an infinite number of independent backscatterers per resolution unit is assumed, commonly referred to as homogeneous regions.

In this context, we present a set of three novel test statistics that aim to distinguish between homogeneous and non-homogeneous returns, particularly between gamma and \(\mathcal{G}^0\) distributed data, assuming the number of looks is known. 
We use properties such as entropy and coefficient of variation.

Entropy is a fundamental concept in information theory with far-reaching applications in pattern recognition, statistical physics, image processing, edge detection and SAR image analysis&nbsp;\cite{Presse2013,MohammadDjafari2015,Avval2021, Nascimento2014,Nascimento2019}. 
Shannon introduced it in 1948&nbsp;\cite{Shannon1948} for a random variable to measure information and uncertainty.
Shannon entropy is a crucial descriptive parameter in statistics, especially for evaluating data dispersion and performing tests for normality, exponentiality and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation is challenging, especially when the model is unknown. In these cases, non-parametric methods are used.
Spacing methods have been discussed as a non-parametric approach in Refs.&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. 
This strategy is flexible and robust because it does not enforce a model or parametric constraints.

The coefficient of variation (CV), introduced in 1896 by Pearson&nbsp;\cite{Pearson1896}, is a relative dispersion measure widely used in various fields of applied statistics, including sampling, biostatistics, medical and biological research, climatology and other fields&nbsp;\cite{hendricks1936sampling,Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is particularly valuable for relating variables with different units. 
This is because when the primary purpose is to compare the variations of several variables, the standard deviation can only serve as an adequate measure of variation if all variables are expressed in the same unit of measurement and have identical means. 
If these conditions are not met, then the CV is the relative measure that is usually used in real applications. 
The variable with the highest CV value has the largest relative dispersion around the mean value&nbsp;\cite{Banik2011}.
The coefficient of variation is the primary measure of heterogeneity in SAR data&nbsp;\cite{Ulaby1986,Touzi1988}.
We study two ways of estimating the coefficient of variation.

The other parameter we study is the Shannon entropy.
Different roughness levels materialized as models for SAR data, have different entropy values, but this fundamental quantity can also be estimated in a model-agnostic way.
We exploit this property and design a bootstrap-improved non-parametric estimator for the Shannon entropy.

We devise test statistics based on these three estimators: 
the classical coefficient of variation, 
a robust version, 
and the Shannon entropy estimator.
We apply these test statistics to generate maps of evidence of homogeneity
that reveal different types of targets in the SAR data. 
We show that our proposed method is superior to existing approaches with simulated data and SAR images.

The article is structured as follows: Section&nbsp;\ref{sec:Background}
deals with statistical modeling and entropy estimation for intensity SAR
data. 
Section&nbsp;\ref{sec:test} outlines hypothesis tests based on non-parametric entropy and coefficients of variation estimators. 
In Section&nbsp;\ref{sec:Results} we present experimental results. 
Finally, we draw conclusions in Section&nbsp;\ref{sec:conclusion}.

@@

The primary models for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully-developed speckle and is a limiting case of the second model. This is interesting due to its versatility in accurately representing regions with different roughness properties&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density
functions (pdfs): 
\begin{align}
    f_Z(z;L, \mu\mid \Gamma_{\text{SAR}})&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z)\label{E:gamma1}\\
    \intertext{ and }
    f_Z(z; \alpha, \gamma, L \mid \mathcal{G}_I^0)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} 
where \(\mu > 0\) is the mean, 
\(\gamma > 0\) is the scale, 
\(\alpha < 0\) measures the roughness, 
\(L \geq 1\) is the number of looks, 
\(\Gamma(\cdot)\) is
the gamma function, 
and \(\mathbbm 1_{A}(z)\) is the indicator function
of the set \(A\).

@@

Vasicek&nbsp;\cite{vasicek1976test} introduced one of the first non-parametric estimators based on spacings. 
Under the assumption that \(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) is a random sample from the distribution \(F(z)\), the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, 
$Z_{(1)}\leq Z_{(2)}\leq\dots\leq Z_{(n)}$ are the order statistics,
and $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing, in which $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

@@

* Correa&nbsp;\cite{correa1995new}:
\begin{equation}
\widehat{H}_{\text{C}}(\bm{Z})=-\frac{1}{n} \sum_{i=1}^n \log \frac{\sum_{j=i-m}^{i+m}(j-i)\left(Z_{(j)}-\widebar{Z}_{(i)}\right)}{n\sum_{j=i-m}^{i+m}\left(Z_{(j)}-\widebar{Z}_{(i)}\right)^2},
\end{equation}
where  $\widebar{Z}_{(i)}=(2 m+1)^{-1} \sum_{j=i-m}^{i+m} Z_{(j)}$, $m< \frac{n}{2}$, $Z_{(i)}=Z_{(1)}$ for $i<1$ and $Z_{(i)}=Z_{(n)}$ for $i>n$. Based on simulations, he showed that his estimator has a smaller mean square error than Vasicek's approach.

@@

These estimators are asymptotically consistent. 
However, we will use improved bootstrap-improved versions because we need them to perform well with small samples.

@@

We aim to test the following hypotheses: 
$$
\begin{cases}
  \mathcal{H}_0: \text{ The data come from the } \Gamma_{\text{SAR}}\text{ law},\\ 
  \mathcal{H}_1:\text{ The data come from the } G_I^0 \text{ distribution}.
\end{cases}
$$
We are testing the hypothesis that the data are fully-developed speckle versus the alternative of data with roughness.
As for the parametric problem, once it is not possible to define the hypothesis $\mathcal{H}_0=\alpha=-\infty$, it is impossible to solve this problem with parametric inference alternatives (such as likelihood ratio, score, gradient, and Wald hypothesis test).
The proposed tests to solve this physical problem in SAR systems are described below.

@@

This test statistic aims to assess the behavior of the data under the null hypothesis using the empirical distribution.
If the data represent fully-developed speckle, the density should center around zero, i.e., $S(\bm{Z};L)\approx 0$. Otherwise, the empirical distribution would shift from zero under the alternative hypothesis, suggesting significant differences and heterogeneous clutter.

@@

The processing time is an important feature, especially considering the application of these estimators to large datasets of SAR images, as seen in Section&nbsp;\ref{sec:Results}.

@@

As visible from Table&nbsp;\ref{tab:table_time}, the accuracy of the test results across the three estimators shows similarities in specific sample sizes. 
However, practical scenarios in SAR image processing often involve small sample sizes, typically obtained over windows of size $7\times7$.  
It is also noteworthy that the \(\widetilde{H}_{\text{AO}}\) estimator exhibited the shortest processing time, followed by \(\widetilde{H}_{\text{E}}\) and \(\widetilde{H}_{\text{C}}\). Considering this aspect, we select the \(\widetilde{H}_{\text{AO}}\) estimator for subsequent simulations. Henceforth,  the test statistical &nbsp;\eqref{Eq:test_e} will be denoted as: \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\).

We now verify the normality of the data generated by the \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\) test.
Figure&nbsp;\ref{fig:Plot_density} shows the empirical densities obtained by applying  the $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$ test to different sample sizes drawn from the $\Gamma_{\text{SAR}}$ distribution, where $L$ takes values $\left\{3,5, 8,11\right\}$ and $\mu=1$. Additionally, Table&nbsp;\ref{tab:table_stat_combined} summarizes the main descriptive statistics, including mean, 
standard deviation&nbsp;(SD), 
variance&nbsp;(Var), 
skewness&nbsp;(SK), 
excessive kurtosis&nbsp;(EK) and 
Anderson--Darling \(p\) values for normality. 
Results with \(p\) values greater than \(0.05\) do not indicate a violation of the normality assumption.
A low variance, skewness, and excessive kurtosis of almost zero indicate limited dispersion, asymmetry, and a light tail.
Normal QQ plots confirm no evidence against a normal distribution, as shown in Figure&nbsp;\ref{fig:Plot_normality_qq}.

@@

After checking the data's normality, we examined the proposed test's abilities in terms of size and power.
Under \(\mathcal{H}_0\), the distribution of the test statistic is asymptotically normal.
Therefore, the \(p\) values are calculated as
\(2\Phi(-|\varepsilon|)\), where \(\Phi\) is the standard
Gaussian cumulative distribution function, and \(\varepsilon\) is the standardized test statistic given by:
\[
\varepsilon=\frac{\widetilde{H}_{\text{AO}}(\bm{Z})-\left[H_{\Gamma_{\text{SAR}}}(L)+\ln \widebar{\bm{Z}}\right]}{\widehat\sigma}.
\]

We have nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and
\SI{10}{\percent}. 
In terms of size, \(1000\) simulations were used for different sample sizes from the $\Gamma_{\text{SAR}}$ distribution, with varying values of $L$, and  $\mu=1$. 
In all cases, the nominal level was achieved. 
We assessed the test power using \(1000\) simulations for different sample sizes from the $G_I^0$ distribution, with  $\mu=1$, and $\alpha=-2$.
The power generally improves with increasing sample size and number of looks.
The results are shown in Table&nbsp;\ref{tab:table_size_power}.

@@

In addition to the \(S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)\) test, we also propose a test statistic based on the classical CV. This test statistic is defined as follows:
\begin{align}
	T_{\text{CV}}=\frac{S}{\widebar{Z}},
\end{align}
where $S$ and $\widebar{Z}$ are the sample standard deviation and mean, respectively.

@@

We used the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) to select the best-fitting distribution.

The AIC deals with the trade-off between the goodness-of-fit and the model's simplicity in terms of the number of model parameters&nbsp;\cite{Burnham2004}. 
The model or distribution with the lowest value of AIC is chosen to be the best.
The BIC assesses goodness-of-fit of a distribution or model, but avoids overfitting by penalising additional degrees of freedom&nbsp;\cite{Dziak2019}. 
The model with the lowest BIC value is chosen as the best.

The AIC and BIC results in Tables&nbsp;\ref{tab:table_aic_gamma}--\ref{tab:table_aic_gio_MnADmedian} indicate that the CV and \(\text{CV}_{\text{MnAD}}\) data from different distributed $\Gamma_{\text{SAR}}$ and \(G_I^0\) synthetic sample sizes match the properties of an LN distribution.
It is important to note that this conclusion was drawn empirically based on a dictionary of analytically tractable distributions and well-defined under biparametric, unimodal, asymmetric, and positive distributions.

Figures&nbsp;\ref{fig:Plot_cv}--\ref{fig:Plot_madmed_gi0_MnADmedian} show empirical and fitted density plots, Q-Q plots, P-P plots, as well as empirical and fitted cumulative distribution functions.
They provide qualitative sources confirming that the LN distribution is the most appropriate.

@@

This section presents the simulations we performed to evaluate the proposed test statistics' performance, followed by applications to SAR data.

@@

The $\alpha$ parameter of the $G_I^0$ distribution is essential for interpreting texture characteristics. 
Values near zero greater than $-3$ suggest extremely textured targets, such as urban zones&nbsp;\cite{Frery2019a}. 
As the value decreases, it indicates regions with moderate texture (in the $\left[-6,-3\right]$ region), related to forest zones, while values below $-6$ correspond to textureless regions, such as pasture, agricultural fields, and water bodies&nbsp;\cite{Neto2023}.

@@

We evaluated the proposed test statistics using three SAR images: one of the coast of Jalisco, Mexico (with a spatial resolution of 20 m both along azimuth and range directions) and two of Illinois, USA (with a spatial resolution of \SI{10}{\meter} both along azimuth and range directions),  acquired by the Sentinel-1B satellite operating in C-band, with VV polarization and intensity format.
The first two images have a size of \(512 \times 512\) pixels, while the third has \(1024 \times 1024\) pixels, and they contain mountainous areas, agricultural regions, water bodies, and urban areas, as shown in  Figures&nbsp;\ref{fig:real_SAR_Images_coe}(a)--(c).

@@

In Figures&nbsp;\ref{fig:Mexico_crops_0.05}, \ref{fig:lake_0.05} and&nbsp;\ref{fig:Illinois_crops_0.05}, the maps of $p$-values composed of a linear gradient of black and white colors, represent the decisions at a \SI{5}{\percent} significance level. 
Dark areas represent values below  $0.05$, indicating evidence to reject the null hypothesis and suggesting heterogeneity in these regions. 
In contrast, values above 0.05 are represented as white areas, indicating no evidence to reject the fully-developed speckle hypothesis.
